---
title: "Knapsack package Group 14"
author: "Raymond Sseguya, Martin Smelik, Vinay"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The following 3 alogorithms based on different computational complexity were implemented to find the solution for the Knapsack problem
1. Brute force search
2. Dynamic programming
3. Greedy heuristic


```{r include=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

set.seed(42)
n <- 1000000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )

```

## Algorithm 1: Brute force search (Computational complexity: O(2^n))

```{r include=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

library(parallel)

brute_force_knapsack <- function(x, W, parallel=FALSE)
{
  stopifnot(is.data.frame(x) == T)
  stopifnot(ncol(x)==2)
  stopifnot(colnames(x)== list("w","v"))
  stopifnot(all(x$v>0))
  stopifnot(all(x$w>0))
  stopifnot(W>=0)
  ideal_value <- 0
  n = nrow(x)
  x$label <- 1:n
  x <- x[which(x$w < W), ]
  n = nrow(x)
  if (parallel == TRUE) {
    cluster = makeCluster(makeCluster(detectCores() - 1))
    permutation <- parSapply(cluster, 1:(2^n-1), function(i) {intToBits(i)[1:n]}, simplify="array")
    weight <- parSapply(cluster, 1:(2^n-1), function(i) {sum(as.numeric(permutation[,i])*x$w) }, simplify="array")
    value <- parSapply(cluster, 1:(2^n-1), function(i) {sum(as.numeric(permutation[,i])*x$v) }, simplify="array")
    stopCluster(cluster)
  }
  else
  {
  permutation <- sapply(1:(2^n-1), function(i) { intToBits(i)[1:n]}, simplify="array")
  weight <- sapply(1:(2^n-1), function(i) { sum(as.numeric(permutation[,i])*x$w) }, simplify="array")
  value <- sapply(1:(2^n-1), function(i) { sum(as.numeric(permutation[,i])*x$v) }, simplify="array")
  }
  for (i in 1:(2^n-1))  
  {
    if (ideal_value<value[[i]] && weight[[i]] < W)
    {
      ideal_value <- value[[i]]
      elements <- as.numeric(permutation[,i])*x$label
    }
  }
  return(list(value=round(ideal_value,0), elements=elements[which(elements>0)])) 
  
}


```

We create a function called brute_force_knapsack with argument x which is a matrix containing 2 rows, weights and its corresponding values. Argument W is the Knapsack capacity. The cumulative value of all possible combinations are computed one after the other and the best combination of elements, i.e. the one with maximum cumulative value whose cumulative weights is within the specified Knapsack weight capacity, are filled in the Knapsack. The function returns the cumulative value of the corresponding elements filled in the knapsack and the element(id) itself.

Time taken to run the algoritm for n= 16 objects : 

```{r message=FALSE, warning=FALSE}
system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 2000))

system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500))

```


##  (*) Parallelize brute force search


The performance gain before parallelizing:

```{r message=FALSE, warning=FALSE, eval= FALSE}

system.time(brute_force_knapsack(x = knapsack_objects[1:30,], W = 3000)) 

```

```{r}

#user    system   elapsed 
#11.72     0.05     12.08 
  

```

The performance gain after parallelizing:

```{r message=FALSE, warning=FALSE, eval = FALSE}

system.time(brute_force_knapsack(x = knapsack_objects[1:30,], W = 3000, parallel = TRUE))

```


```{r}

#user    system   elapsed 
#1.94     1.24     11.79 
  

```

   

## Algorithm 2: Dynamic programming (Computational complexity: O(Wn))

```{r include=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

knapsack_dynamic <- function(x,W)
{
  stopifnot(is.data.frame(x) == T)
  stopifnot(ncol(x)==2)
  stopifnot(colnames(x)== list("w","v"))
  stopifnot(all(x$v>0))
  stopifnot(all(x$w>0))
  stopifnot(W>=0)
  n = nrow(x)
  x$ord <- 1:n
  x <- x[which(x$w < W), ]
  n = nrow(x)
  m <- matrix(0, nrow=n, ncol=W+1)
  path <- as.data.frame(matrix(0,nrow = n, ncol = W+1))

  for(i in 1:n)
  {
    for(j in 1:W+1)
    {
      if (x$w[[i]] >j-1 && i==1) {path[[1,j]] <- list(0)}
      else if (x$w[[i]] >j-1) 
      {
        m[[i,j]] <- m[[i-1,j]]
        path[[i,j]] <- path[[i-1,j]]
      }
      else 
      {
        if(i==1) 
        {
          m[[i,j]] <- x$v[[i]]
          path[[i,j]] <- list(x$ord[[i]])
        }
        else
        {
          m[[i,j]] <- max(m[[i-1,j]],m[[i-1,j-x$w[[i]]]] + x$v[[i]])
          if(m[[i-1,j]] < m[[i-1,j-x$w[[i]]]] + x$v[[i]])
            { path[[i,j]] <- rlist::list.append(path[[i-1,j-x$w[[i]]]],x$ord[[i]]) }
          else
            path[[i,j]] <- path[[i-1,j]]
        }
      }
    }
  }
  elements <- path[[n,W+1]]
  elements = elements[which(elements>0)]
  return(list(value= round(m[[n,W+1]],0), elements=unlist(elements, use.names = FALSE) ))
}

```

Tabular method is used to solve the problem using Dynamic programming method. 

Time taken to run the algoritm for n= 500 objects :

```{r message=FALSE, warning=FALSE, eval=FALSE}
system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 2000))

system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500))

```



```{r}
#For W=2000
#user    system   elapsed 
#84.57     0.02     84.69 
  
#For W=3500

#user    system   elapsed 
#391.42     0.16    392.93 
```


## Algorithm 3: Greedy heuristic (Computational complexity: O(nlogn))

```{r include=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

greedy_knapsack <- function(x,W)
{
  stopifnot(is.data.frame(x) == T)
  stopifnot(ncol(x)==2)
  stopifnot(colnames(x)== list("w","v"))
  stopifnot(all(x$v>0))
  stopifnot(all(x$w>0))
  stopifnot(W>=0)
  n = nrow(x)
  x$ord <- 1:n
  x <- x[which(x$w < W), ]
  n = nrow(x)
  x$z <- x$v/x$w
  x <-x [with(x, order(-z)),]
  i<- 1
  weight <- 0
  value <- 0
  elements <- list()
  while (i<n+1 && weight + x$w[[i]]<W  )
  {
    weight <- weight + x$w[[i]] 
    value <- value + x$v[[i]]
    elements[[i]] <- x$ord[[i]]
    i<- i+1
  }

  return(list(value=round(value,0), elements=unlist(elements, use.names = FALSE) ))
}

```

The ratios of each value to its corresponding weights are considered to fill up the knapsack. The elements with the higher ratio is put into the knapsack first. 

Time taken to run the algoritm for n= 1000000 objects :

```{r message=FALSE, warning=FALSE, eval=FALSE}
system.time(greedy_knapsack(x = knapsack_objects[1: 1000000,], W = 2000))

system.time(greedy_knapsack(x = knapsack_objects[1: 1000000,], W = 3500))

```


```{r}
#For W=2000
#user    system   elapsed 
#0.18      0.01      0.20 
   
#For W=3500
#user    system   elapsed 
#0.36      0.00      0.38
```   

## 1.1.5 Implement a test suite





```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# context("brute_force_knapsack")
# 
# set.seed(42)
# n <- 2000
# knapsack_objects <- data.frame(
#   w=sample(1:4000, size = n, replace = TRUE),
#   v=runif(n = n, 0, 10000)
# )
# 
# test_that("Correct object is returned", {
#   expect_silent(bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500))
#   expect_named(bfk, c("value", "elements"))
# })
# 
# 
# test_that("functions rejects errounous input.", {
#   expect_error(brute_force_knapsack("hej", 3500))
#   expect_error(brute_force_knapsack(x = knapsack_objects[1:8,], W = -3500))
# })
# 
# test_that("Function return correct results.", {
#   bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
#   expect_equal(round(bfk$value), 16770)
#   expect_true(all(round(bfk$elements) %in% c(5, 8)))
#   
#   bfk <- brute_force_knapsack(x = knapsack_objects[1:12,], W = 3500)
#   expect_equal(round(bfk$value), 16770)
#   expect_true(all(round(bfk$elements) %in% c(5, 8)))
#   
#   bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 2000)
#   expect_equal(round(bfk$value), 15428)
#   expect_true(all(round(bfk$elements) %in% c(3, 8)))
#   
#   bfk <- brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000)
#   expect_equal(round(bfk$value), 15428)
#   expect_true(all(round(bfk$elements) %in% c(3, 8)))
#   
#   st <- system.time(bfk <- brute_force_knapsack(x = knapsack_objects[1:16,], W = 2000))
#   expect_true(as.numeric(st)[2] >= 0.00)
# })

```




```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# context("greedy_knapsack")
# 
# set.seed(42)
# n <- 2000
# knapsack_objects <- data.frame(
#   w=sample(1:4000, size = n, replace = TRUE),
#   v=runif(n = n, 0, 10000)
# )
# 
# test_that("Correct object is returned", {
#   expect_silent(gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 3500))
#   expect_named(gk, c("value", "elements"))
# })
# 
# test_that("functions rejects errounous input.", {
#   expect_error(greedy_knapsack("hej", 3500))
#   expect_error(greedy_knapsack(x = knapsack_objects[1:8,], W = -3500))
# })
# 
# test_that("Function return correct results.", {
#   gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 3500)
#   expect_equal(round(gk$value), 15428)
#   expect_true(all(round(gk$elements) %in% c(3, 8)))
#   
#   gk <- greedy_knapsack(x = knapsack_objects[1:12,], W = 3500)
#   expect_equal(round(gk$value), 15428)
#   expect_true(all(round(gk$elements) %in% c(3, 8)))
#   
#   gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 2000)
#   expect_equal(round(gk$value), 15428)
#   expect_true(all(round(gk$elements) %in% c(3, 8)))
#   
#   gk <- greedy_knapsack(x = knapsack_objects[1:12,], W = 2000)
#   expect_equal(round(gk$value), 15428)
#   expect_true(all(round(gk$elements) %in% c(3, 8)))
#   
#   st <- system.time(gk <- greedy_knapsack(x = knapsack_objects[1:16,], W = 2000))
#   expect_true(as.numeric(st)[2] <= 0.01)
#   
#   gk <- greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
#   expect_equal(round(gk$value), 192647)
#   
#   gk <- greedy_knapsack(x = knapsack_objects[1:1200,], W = 3500)
#   expect_equal(round(gk$value), 270290)
# })

```




```{r  eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# context("knapsack_dynamic")
# 
# 
# set.seed(42)
# n <- 2000
# knapsack_objects <- data.frame(
#   w=sample(1:4000, size = n, replace = TRUE),
#   v=runif(n = n, 0, 10000)
# )
# 
# test_that("Correct object is returned", {
#   expect_silent(bfk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500))
#   expect_named(bfk, c("value", "elements"))
# })
# 
# 
# test_that("functions rejects errounous input.", {
#   expect_error(knapsack_dynamic("hej", 3500))
#   expect_error(knapsack_dynamic(x = knapsack_objects[1:8,], W = -3500))
# })
# 
# test_that("Function return correct results.", {
#   bfk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
#   expect_equal(round(bfk$value), 16770)
#   expect_true(all(round(bfk$elements) %in% c(5, 8)))
#   
#   bfk <- knapsack_dynamic(x = knapsack_objects[1:12,], W = 3500)
#   expect_equal(round(bfk$value), 16770)
#   expect_true(all(round(bfk$elements) %in% c(5, 8)))
#   
#   bfk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 2000)
#   expect_equal(round(bfk$value), 15428)
#   expect_true(all(round(bfk$elements) %in% c(3, 8)))
#   
#   bfk <- knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
#   expect_equal(round(bfk$value), 15428)
#   expect_true(all(round(bfk$elements) %in% c(3, 8)))
#   
#   st <- system.time(bfk <- knapsack_dynamic(x = knapsack_objects[1:16,], W = 2000))
#   expect_true(as.numeric(st)[2] >= 0.00)
# })

```

We implemented the test suit also for dynamic approach and found out that the greedy algorithm returns the same solution as in test suit.



## 1.1.6 Profile code

To improve our code, we decided to check and exclude all the objects which weight was greater than W in all the algorithms.

## 1.1.8 Parallelizing

We included an option to parallelize the algorithm and the results can be seen above. It obviously helped especially for greater n